---
title: "Consequences of Covid-19 for Candle Critiques"
author: "Greg Rousis"
date: "`r format(Sys.Date(), '%b %d %Y')`"
output: 
  github_document: default
  html_document:
    theme: readable
    highlight: haddock
    code folding: show
    toc: true
    toc_float: true
    toc_depth: 3
---


## The data

There are 2 datasets in this report: one for people's reviews of scented candles and one for reviews of unscented candles. The datasets have the following variables:

  1. `CandleID`: A numeric id variable for candle 1-3 in each category.
  2. `Date`: The date of the review.
  3. `Rating`: The 1-5 rating for the review.
  4. `Review`: For the scented candles, the text of the review.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', fig.width = 12, fig.height = 8)

library(tidyverse)
library(readxl)
library(patchwork)
theme_set(theme_classic(base_size = 18))

candles_scented <- read_excel("Scented_all.xlsx")
candles_unscented <- read_excel("Unscented_all.xlsx")

```


## Part 1: Average daily ratings for scented and unscented candles

Since the onset of COVID, average ratings for scented candles have dropped preciptiously, whereas those for unscented candles have remained relatively stable.

```{r separate plots}

p1 <- candles_scented %>% 
  filter(Date >= "2017-01-01") %>% 
  group_by(Date) %>% 
  mutate(Rating_Avg = mean(Rating)) %>%
  distinct(Date,
           .keep_all = TRUE) %>% 
  ggplot() +
  aes(x = Date,
      y = Rating_Avg) +
  geom_point(shape = 1, 
             alpha = 0.3) +
  geom_vline(xintercept = as.POSIXct(as.Date("2020-01-20")),
             linetype = 4) +
  geom_smooth(method = "loess",
              color = "goldenrod4",
              fill = "thistle1") +
  scale_x_datetime(date_breaks = "6 months",
                   date_labels = "%m-%Y") +
  ylab(label = "Average Daily Rating")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle(label = "Average Daily Ratings\n of Scented Candles") +
  theme(plot.title = element_text(hjust = 0.5))


p2 <- candles_unscented %>% 
  filter(Date >= "2017-01-01") %>% 
  group_by(Date) %>% 
  mutate(Rating_Avg = mean(Rating)) %>%
  distinct(Date,
           .keep_all = TRUE) %>% 
  ggplot() +
  aes(x = Date,
      y = Rating_Avg) +
  geom_vline(xintercept = as.numeric(as.Date("2020-01-20"))) +
  geom_vline(xintercept = as.POSIXct(as.Date("2020-01-20")),
             linetype = 4) +
  geom_point(shape = 1,
             alpha = 0.3) +
  geom_smooth(method = "loess",
              color = "goldenrod4",
              fill = "thistle1") +
  scale_x_datetime(date_breaks = "6 months",
                   date_labels = "%m-%Y") +
  scale_fill_brewer() +
  ylab(label = "Average Daily Rating")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle(label = "Average Daily Ratings\n of Unscented Candles") +
  theme(plot.title = element_text(hjust = 0.5))



p1 + p2

```


## Part 2: Combined plot



```{r combine datasets, eval=FALSE}

candles_full <- bind_rows(
	scented = candles_scented, 
	unscented = candles_unscented, 
	.id = "Scented"
)

```

```{r}

candles_full <- bind_rows(
	scented = candles_scented, 
	unscented = candles_unscented, 
	.id = "Scented"
)

full_filtered <- candles_full %>% 
  filter(Date >= "2017-01-01") %>% 
  group_by(Date, Scented) %>% 
  mutate(Rating_Avg = mean(Rating)) %>%
  distinct(Date,
           .keep_all = TRUE)

ggplot(full_filtered) +
  aes(x = Date,
      y = Rating_Avg,
      color = Scented,
      group = Scented,
      linetype = Scented) +
  geom_vline(xintercept = as.POSIXct(as.Date("2020-01-20")),
             linetype = 4) +
  geom_point(shape = 1,
             alpha = 0.5)+
  geom_smooth(method = "loess",
              se = FALSE) +
  ylab(label = "Average Daily Rating") +
  ggtitle(label = "Average Daily Ratings\n of Scented and Unscented Candles") +
  theme(plot.title = element_text(hjust = 0.5))

```

Which of these two sets of plots do you find easier to interpret?

I think both tell the story that scented candles started performing worse once COVID and its attendant olfactory effects kicked in, athough it's easier to see the relative trend when both scented and unscented reviews are in the same graph.


## Part 3: Proportion of reviews mentioning "no scent"

Looking at the content of people's reviews, it appears that a lot of complaints came in the winter months of 2019 (cold and flu season), but are much more consistent in 2020, likely reflecting the impact of COVID. Furtermore, the trend if people complaining that they can't smell their candles appears to mirror the occurrence of allergy season, cold and flu season, and then the onset of COVID.

It also looks like the proportion of people complaining that their candles don't smell is larger in 2020 moths relative to 2019.


```{r no scent function, include=FALSE}
str_no_scent <- function(x){
  as.numeric(
  	case_when(
  		str_detect(x, "[Nn]o scent") ~ "1", 
  		str_detect(x, "[Nn]o smell") ~ "1",
  		str_detect(x, "[Dd]oes not smell like") ~ "1",
  		str_detect(x, "[Dd]oesn't smell like") ~ "1",
  		str_detect(x, "[Cc]an't smell") ~ "1",
  		str_detect(x, "[Cc]annot smell") ~ "1",
  		str_detect(x, "[Ff]aint smell") ~ "1",
  		str_detect(x, "[Ff]aint scent") ~ "1",
  		str_detect(x, "[Dd]on't smell") ~ "1",
  		str_detect(x, "[Ll]ike nothing") ~ "1",
  		TRUE ~ "0"
  	)
  )
}
```

```{r}

candles_scented %>% 
  filter(Date >= "2019-01-01", Date < "2021-01-01") %>% 
  mutate(Month = lubridate::floor_date(Date, "month")) %>% 
  mutate(noscent = str_no_scent(Review)) %>% 
  group_by(Month, noscent) %>% 
  summarize(n = n()) %>% 
  mutate(prop = n/sum(n)) %>%
  mutate(Year = as.factor(ifelse(Month > "2019-12-01", 2020, 2019))) %>% 
  filter(noscent == 1) %>% 
  ggplot() +
  aes(x = Month,
      y = prop,
      fill = Year) +
  geom_bar(stat = "identity") +
  scale_x_datetime(date_breaks = "1 month",
                   date_labels = "%m-%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(label = "Propotion of reviews mentioning no scent")




```

