---
title: "hw03"
author: "Greg Rousis"
date: "2/18/2021"
output: 
  html_document: default
  github_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#loading required packages

setwd("C:/Users/Greg Rousis/GitHub/hw-gregrousis/hw03")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2)
theme_set(theme_classic())

```

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r data}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")

print(dictionary, n = 36)

```

# Question 1: inspect item responses

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
     
```{r data inspection}
green_data %>% 
  summary()
```

## There are some -99's, which need to be recoded as missing

```{r data inspection}

green_data <- green_data %>% 
  mutate_at(vars(comp1:student),
            ~ifelse(. == -99, NA, .)
  )

green_data %>% 
  summary()

green_data <- 
  green_data %>% 
  mutate(student = factor(student, levels = c(1, 2))) %>% 
  mutate(
    student = recode(student, "1" = "Not a student", "2" = "Student")
  )

  
  
melt.green <- melt(green_data)

head(melt.green)

ggplot(melt.green) +
  aes(x = value) +
  geom_histogram() +
  facet_wrap(~variable, 
             scales = "free")
```


# Question 2: compute total scores

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
     
```{r total scores}

# pull the reverse-coded items from the dictionary
reversed <- dictionary %>%
  filter(Keying <= 0) %>%
  pull(Item)

#recode to match direction of other items

green_data <- green_data %>%
  mutate(
    across(all_of(reversed),
           ~  6-.x,
           .names = "{.col}r")
  )

green_data %>% 
  select(green2, green2r)

# compute total scores

green_data <- 
  green_data %>% 
  rowwise() %>% 
  mutate(green_tot = mean(c(green1, green2r, green3, green4, green5), na.rm = TRUE),
         comp_tot = mean(c(comp1, comp2, comp3, comp4, comp5, comp6r, comp7r, comp8r, comp9r, comp10r), 
                         na.rm = TRUE),
         intel_tot = mean(c(intel1, intel2, intel3, intel4, intel5, intel6, intel7r, intel8r, intel9r, intel10r),
                          na.rm = TRUE),
         open_tot = mean(c(open1, open2, open3, open4, open5, open6, open7r, open8r, open9r, open10r), 
                         na.rm = TRUE)) %>% 
  ungroup()
         


```

# Question 3: Rescaling using POMP

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
     
```{r pomp}

# computing pomp scores (100*(raw-min)/(max-min))

green_data <- 
  green_data %>%
  rowwise() %>% 
  mutate(green_pomp = (100*(green_tot - 1))/(5-1), na.rm = TRUE,
         comp_pomp = (100*(comp_tot - 1))/(5-1), na.rm = TRUE,
         intel_pomp = (100*(intel_tot - 1))/(5-1), na.rm = TRUE,
         open_pomp = (100*(open_tot - 1))/(5-1), na.rm = TRUE) %>% 
  ungroup()


```


# Question 4: distribution plots

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
```{r pomp plots}

ggplot(green_data) +
  aes(x = green_pomp) +
  geom_bar()

ggplot(green_data) +
  aes(x = comp_pomp) +
  geom_histogram(binwidth = 1.5)

ggplot(green_data) +
  aes(x = intel_pomp) +
  geom_bar()

ggplot(green_data) +
  aes(x = open_pomp) +
  geom_bar()


```

# Question 5: scatterplots

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
     
```{r scatters}

green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(x = comp_pomp,
      y = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")
  

green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(x = intel_pomp,
      y = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")

  
green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(x = open_pomp,
      y = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")

```

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
