---
title: "hw03"
author: "Greg Rousis"
date: "2/18/2021"
output: 
  html_document: default
  github_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#loading required packages

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2)
theme_set(theme_classic())

```

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r data}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")

print(dictionary, n = 36)

```

#Question 1:

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
     
```{r data inspection}
green_data %>% 
  summary()
```

## There are some -99's, which need to be recoded as missing

```{r data inspection}

green_data <- green_data %>% 
  mutate_at(vars(comp1:student),
            ~ifelse(. == -99, NA, .)
  )

green_data %>% 
  summary()
  
melt.green <- melt(green_data)

head(melt.green)

ggplot(melt.green) +
  aes(x = value) +
  geom_histogram() +
  facet_wrap(~variable, 
             scales = "free")
```


# Question 2:

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
     
```{r total scores}

# pull the reverse-coded items from the dictionary
reversed <- dictionary %>%
  filter(Keying <= 0) %>%
  pull(Item)

#recode to match direction of other items

green_data <- green_data %>%
  mutate(
    across(all_of(reversed),
           ~  6-.x,
           .names = "{.col}r")
  )

green_data %>% 
  select(green2, green2r)

```

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
