---
title: "hw03"
author: "Greg Rousis"
date: "2/18/2021"
output: 
  github_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#loading required packages

setwd("C:/Users/Greg Rousis/GitHub/hw-gregrousis/hw03") # BMW: Never, ever use setwd() ;)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2) # BMW: I recommend against reshape2. It's retired and no longer maintained. Use tidyr::pivot_longer() etc. instead. Especially don't use reshpae2 and tidyverse in the same script.
library(gridExtra)
library(kableExtra)
theme_set(theme_classic())

```

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r data}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")

print(dictionary, n = 36)

```



# Question 1: inspect item responses

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
     
```{r data inspection}
green_data %>% 
  summary()
```


### There are some -99's, which need to be recoded as missing

```{r missing data}

green_data <- green_data %>% 
  mutate(across(comp1:student, # BMW: mutate_at() is retired. Prefer across() instead.
            ~ifelse(. == -99, NA, .))
  )

green_data %>% 
  summary()

green_data <- 
  green_data %>% 
  mutate(student = factor(student, levels = c(1, 2), labels = c("Not a student", "Student"))) # BMW: A somewhat simpler approach

  
melt.green <- melt(green_data) # BMW: As above, use pivot_longer() instead. 

head(melt.green) # BMW: Don't include "check" views like this in your final script. 

ggplot(melt.green) +
  aes(x = value) +
  geom_histogram() +
  facet_wrap(~variable, 
             scales = "free")
```



# Question 2: compute total scores

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
     
```{r total scores}

# pull the reverse-coded items from the dictionary
reversed <- dictionary %>%
  filter(Keying <= 0) %>%
  pull(Item)

#recode to match direction of other items

green_data <- green_data %>%
  mutate(
    across(all_of(reversed),
           ~  6-.x,
           .names = "{.col}r")
  )

green_data %>% 
  select(green2, green2r)

# compute total scores

green_data <-
  green_data %>%
  rowwise() %>%
  mutate(
    green_tot = mean(c(green1, green2r, green3, green4, green5), na.rm = TRUE),
    comp_tot = mean(
      c(
        comp1,
        comp2,
        comp3,
        comp4,
        comp5,
        comp6r,
        comp7r,
        comp8r,
        comp9r,
        comp10r
      ),
      na.rm = TRUE
    ),
    intel_tot = mean(
      c(
        intel1,
        intel2,
        intel3,
        intel4,
        intel5,
        intel6,
        intel7r,
        intel8r,
        intel9r,
        intel10r
      ),
      na.rm = TRUE
    ),
    open_tot = mean(
      c(
        open1,
        open2,
        open3,
        open4,
        open5,
        open6,
        open7r,
        open8r,
        open9r,
        open10r
      ),
      na.rm = TRUE
    )
  ) %>%
  ungroup()
```



# Question 3: Rescaling using POMP

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
     
```{r pomp}

# computing pomp scores (100*(raw-min)/(max-min))

green_data <- 
  green_data %>%
  rowwise() %>% # BMW: You don't need rowwise() here. All of the operations you are doing are vectorized. 
  mutate(green_pomp = (100*(green_tot - 1))/(5-1), na.rm = TRUE,
         comp_pomp = (100*(comp_tot - 1))/(5-1), na.rm = TRUE,
         intel_pomp = (100*(intel_tot - 1))/(5-1), na.rm = TRUE,
         open_pomp = (100*(open_tot - 1))/(5-1), na.rm = TRUE) %>% 
  ungroup()
```



# Question 4: distribution plots

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
  **Answer:** Comp and openness seem skewed, and intellectual curiosity has a bimodal distribution. Green reputation
  seems the closest to a normal distribution.
  
```{r pomp plots}

#creating plots to inspect data
green_plot <- 
  ggplot(green_data) +
  aes(x = green_pomp) +
  geom_bar()

comp_plot <- 
  ggplot(green_data) +
  aes(x = comp_pomp) +
  geom_histogram(binwidth = 1.5)

intel_plot <- 
  ggplot(green_data) +
  aes(x = intel_pomp) +
  geom_bar()

open_plot <- 
  ggplot(green_data) +
  aes(x = open_pomp) +
  geom_bar()

#arranging data
grid.arrange(green_plot, comp_plot, intel_plot, open_plot) # BMW: We will look at this next week, but check out the patchwork package for an elegant approach this this. 

```



# Question 5: scatterplots

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
     
  **Answer:** green reputation is positively correlated with all 3 personality variables. However, the association 
  appears to differ between students and non-students. Compassion is positively associated with green reputation among
  both students and non-students, but students appear to report higher mean levels of compassion than non-students. 
  Green reputation is positively associated with intellectual curiosity among students, but seems to  have little 
  association with intellectual curiosity among non-students. Likewise, the association between green repuatation and 
  openness is seemingly stronger among students than non-students. 
     
```{r scatters}

green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = comp_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")
  

green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = intel_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")

  
green_data %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = open_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")

```



# Question 6: rainfall plots
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data
  points).
  
  **answer:** Non-students appear to have a bimodal distribution of green reputation: one mode is around 50 (out of
  100) and another around 75 (out of 100). For students, green reputation resembles a more normal distribution.
  
  
```{r rainfall}

green_data %>% 
  filter(!is.na(student)) %>% 
ggplot() +
  aes(y = student,
      x = green_pomp,
      fill = student,
      color = student) +
  geom_jitter(height = .15) +
  geom_boxplot(color = "black",
               alpha = .5,
               width = .1,
               size = .5) +
  ggdist::stat_slab(height = .3,
                    color = "black",
                    size = .2,
                    alpha = .5,
                    position = position_nudge(y = .2))

```

  
  
# Question 7: summary table

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and
  non-students.
  
```{r table}

tab <- green_data %>% 
  na.omit() %>% 
  group_by(student) %>% 
  summarize(mean_green = mean(green_pomp),
            sd_green = sd(green_pomp),
            median_green = median(green_pomp),
            min_green = min(green_pomp),
            max_green = max(green_pomp),
            mean_comp = mean(comp_pomp),
            sd_comp = sd(comp_pomp),
            median_comp = median(comp_pomp),
            min_comp = min(comp_pomp),
            max_comp = max(comp_pomp),
            mean_intel = mean(intel_pomp),
            sd_intel = sd(intel_pomp),
            median_intel = median(intel_pomp),
            min_intel = min(intel_pomp),
            max_intel = max(intel_pomp),
            mean_open = mean(open_pomp),
            sd_open = sd(open_pomp),
            median_open = median(open_pomp),
            min_open = min(open_pomp),
            max_open = max(open_pomp))

  
kable(tab)

```

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
