---
title: "Beer Advocate: ABV and quality ratings"
author: "Greg Rousis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(here)
```

For this activity, download the data file at this location to the `data` subfolder in your `participation` folder. 
This dataset contains information about different catagories of beer as rated on <beeradvocate.com>.
The data are originally from [here](http://scottjanish.com/relationship-of-abv-to-beer-scores/).

```{r}
url <- "https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/beer-advocate.csv"
```

The dataset contains 3 variables:

  1. Beer category
  2. Average alcohol by volume (ABV) in percentage (%)
  3. Average quality rating (rAVG) on a 1–5 scale
  
You should write a script to do the following:

  1. Download the dataset with an appropriate filename.
  2. Import the dataset into R.
     Be sure to check for any importing problems and correct as needed.
  
  
```{r data}

if ( !dir.exists( here::here("participation", "data") ) ) {
  dir.create( here::here("participation", "data"), recursive = TRUE )
}

beer_data <- read_csv(url,
                      col_types = cols(
                        `Beer Category` = col_character(),
                        `Average Abv` = col_double(),
                        rAVG = col_double()
                      )
)

beer_data$`Average ABV` <- as.double(beer_data$`Average ABV`)

beer_data

write_csv(
  beer_dat,
  here::here("participation", "data", "beer_data.csv")
)
  

```


  3. Make a figure illustrating the relationship between ABV and beer quality.
  
  
```{r plot}

ggplot(beer_data) +
  aes(x = `Average ABV`,
      y = rAVG) +
  geom_jitter() +
  theme_classic()

```

  4. Compute a summary table with the mean and SD of each variable, along with their correlation.
  
```{r summary}

beer_data %>% 
  summarize(across(
    where(is.double),
    list(
      mean = ~mean(.x, na.rm = TRUE),
      sd = ~sd(.x, na.rm = TRUE)),
    .names = "{.col}.{.fn}")
  )


   



list(mean = mean, sd = sd, na.rm = TRUE)


```

  5. Describe the results.
